---
title: "Quickstart: Calculating Survey Weights"
author: "Martin Barner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart: Calculating Survey Weights}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# General workflow

## Prerequisites; You need to have:
- your data as a data.frame
    - one column should uniquely specify the stratum of the record
- your sampling frame as a data.frame
    - one row per stratum
    - one column with the stratum names (corresponding exactly to the values in the stratum name column in the data)
    - one column with the corresponding population numbers

```{r,include=F}
library(surveyweights)
# example data:
mydata<-data.frame(somevar1=c(1:5),
                   somevar2=c("one","more","test","variable",NA),
                   strata_names=c("a","a","b","b","a"),
                   cluster_names=c("x","y","z","z","y"))

# example sampling frame:
mysf<-data.frame(strata=c("a","b"),pops=c(100,1000))
clustersf<-data.frame(strata=c("x","y","z"),pops=c(50,100,200))


```

### 1. Example data & sampling frame:
Example data:
```{r, eval=T}
mydata
```
Example sampling frame:
```{r, eval=T}
mysf
```
Note that the values in  `mydata$strata_names` all appear exactly in `mysf$strata`.

### Calculate weights
First we generate a weighting function that matches our dataset and samplingframe. Then we use that function to calculate weights for our dataset and/or different subsets of it.

#### Generate weight function
To generate a weight function, we need to provide at least:

- the sampling frame
- which column in the data denote the stratum
- which columns in the sampling frame are what

So we can do:
```{r, eval=T}
myweighting<-weighting_fun_from_samplingframe(sampling.frame = mysf,
                                            data.stratum.column = "strata_names",
                                            sampling.frame.population.column = "pops",
                                            sampling.frame.stratum.column = "strata")

```
  
`weighting` are not the weights itself, but a _function to generate weights_.

### Calculate weights:

Now to get the actual weights:

```{r,eval=T}
myweighting(mydata)

```

This is useful because we can calculate weights for subsets on the fly:

```{r}
# remove NA records:
mydata_subset<-mydata[!is.na(mydata$somevar2),]

# there are now only 2 records in stratum "a":
mydata_subset

# so the weights are different:
myweighting(mydata_subset)

```


## Combining multiple sampling frames

Assume that in addition to the above, we have a second sampling frame with populations sizes for example for cluster inside the strata:
```{r}

clustersf
```

As above, we generate a weighting function for the cluster:

```{r}
myweighting_cluster<-weighting_fun_from_samplingframe(sampling.frame = clustersf,
                                            data.stratum.column = "cluster_names",
                                            sampling.frame.population.column = "pops",
                                            sampling.frame.stratum.column = "strata")

```

The weights we get with `myweighting_cluster` are incorrect, because the cluster between strata are not proportional. We can combine the two weighting functions like this:
```{r}

myweighting_combined<-combine_weighting_functions(myweighting,
                                                  myweighting_cluster)

myweighting_combined(mydata)

# in comparison: the cluster weights only:
myweighting_cluster(mydata)

# This of course still works for subsets:
myweighting_cluster(mydata_subset)

```



